Runtime-Assisted Convergence in Replicated Data Types

### 2021-11-20 evaluation.tex
Quark --> Carmot. Related work.

\quark's background merges however pose a new problem as they create
new versions on the local branch in the background while the user is
busy editing an older version. When the user attempts to write their
version of the document to the store, simply committing it would
effectively override the concurrent updates from other users obtained
via background merges. This problem is depicted in
Fig.~\ref{fig:local-merge}. The solution, fortunately, is
straightforward: we \emph{merge} the user-submitted value with the
(value of) the latest version on the local branch to create a new
version that includes the updates from either direction. The LCA for
this merge is simply the last version on the local branch read by the
user. This merge need not be synchronized as it doesn't alter the LCAs
between any two branches. The merged value can then returned
to the user as the result of the write.  Quark's \C{write} is a
function of type $\texttt{Value} \rightarrow \texttt{Value}$, where
$\texttt{Value}$ is \C{doc} in the current application.

@@

\quark system replicates the contents of each branch across all the
replicas as fast as the network allows. However, for a user $A$ to
see the changes made by the other user $B$, the changes have to be
reflected in $A$'s local version, which can only happen through a
merge operation. Since \quark synchronizes merge operations globally,
it induces additional delay before $A$ can see $B$'s changes. We call
this additional delay \emph{staleness} as with progression of time,
$B$'s version known to $A$ becomes increasingly stale. At the
system-level, an increase in staleness effectively delays the
convergence (but doesn't stop it, as proved by
Theorem~\ref{thm:progress}). To understand the staleness behavior in
\quark, we repeat the collaborative editing experiment, this time
measuring the staleness value at every merge. We do this by annotating
every version $v$ with the timestamp $t$ of its creation time. When
$v$ is is merged into a remote branch $b$ at a later time $t'$, the
difference $t' - t$ denotes the staleness of $v$ w.r.t the new version
on $b$. Multiple such staleness measurements are recorded for each
experiment and analyzed to compute the 10th, 50th, and 90th percentile
values. Fig.~\ref{fig:staleness} shows the results as we repeat
experiments with increasing number of concurrent users. While
staleness remains in the order of milliseconds with fewer ($\leq 9$)
concurrent users, it increases super-linearly as we increase the
number of collaborators by 3 until 30. While increased staleness is
inevitable in our approach due to synchronized merges, we believe the
increase can be contained by choosing the order of merges to avoid the
``starvation'' of some branches. Note that relaxing the
linearizability constraint on merges would completely eliminate the
staleness overhead, but the resultant Git-like system fails to
converge due to anomalous executions described in
Sec.~\ref{sec:motivation}\footnote{
  Git admits such anamolous version history graphs where two branches
  can have the same set of commits and yet differ in their final
  version. Appendix describes two such cases we observed on Github}.

Our experiments bring to the fore an inherent tradeoff among the
competing concerns of RDTs, namely (\rom{1}). The ease of achieving
convergence, (\rom{2}) Latency, and (\rom{3}).  Staleness. While CRDTs
try to optimize for latency and staleness, they require a significant
amount of development and verification effort to be expended to ensure
convergence~\cite{kleppmann2017}. In contrast, \quark lets developers
derive convergent-by-construction RDTs from ordinary data data types
that are optimized for latency, but incur a staleness overhead that
delays the time to convergence.

>>>>

\quark's background merges however pose a new problem as they create
new versions on the local branch in the background while the user is
busy editing an older version. When the user attempts to write their
version of the document to the store, simply committing it would
effectively override the concurrent updates from other users obtained
via background merges. This problem is depicted in
Fig.~\ref{fig:local-merge}. The solution, fortunately, is
straightforward: we \emph{merge} the user-submitted value with the
(value of) the latest version on the local branch to create a new
version that includes the updates from either direction. The LCA for
this merge is simply the last version on the local branch read by the
user. This merge need not be synchronized as it doesn't alter the LCAs
between any two branches. The merged value can then returned
to the user as the result of the write.  \quark's \C{write} is a
function of type $\texttt{Value} \rightarrow \texttt{Value}$, where
$\texttt{Value}$ is \C{doc} in the current application.

@@

\quark system replicates the contents of each branch across all the
replicas as fast as the network allows. However, for a user $A$ to
see the changes made by the other user $B$, the changes have to be
reflected in $A$'s local version, which can only happen through a
merge operation. Since \quark synchronizes merge operations globally,
it induces additional delay before $A$ can see $B$'s changes. We call
this additional delay \emph{staleness} as with progression of time,
$B$'s version known to $A$ becomes increasingly stale. At the
system-level, an increase in staleness effectively delays the
convergence (but doesn't stop it, as proved by
Theorem~\ref{thm:progress}). To understand the staleness behavior in
\quark, we repeat the collaborative editing experiment, this time
measuring the staleness value at every merge. We do this by annotating
every version $v$ with the timestamp $t$ of its creation time. When
$v$ is is merged into a remote branch $b$ at a later time $t'$, the
difference $t' - t$ denotes the staleness of $v$ w.r.t the new version
on $b$. Multiple such staleness measurements are recorded for each
experiment and analyzed to compute the 10th, 50th, and 90th percentile
values. Fig.~\ref{fig:staleness} shows the results as we repeat
experiments with increasing number of concurrent users. While
staleness remains in the order of milliseconds with fewer ($\leq 9$)
concurrent users, it increases super-linearly as we increase the
number of collaborators by 3 until 30. While increased staleness is
inevitable in our approach due to synchronized merges, we believe the
increase can be contained by choosing the order of merges to avoid the
``starvation'' of some branches. Note that relaxing the
linearizability constraint on merges would completely eliminate the
staleness overhead, but the resultant Git-like system fails to
converge due to anomalous executions described in
Sec.~\ref{sec:motivation}\footnote{
  Git admits anamolous version history graphs where two branches can
  have the same set of commits and yet differ in their final version.
  Appendix describes two such cases we observed on Github}.

Our experiments bring to the fore an inherent tradeoff among the
competing concerns of RDTs, namely (\rom{1}). The ease of programming
convergence, (\rom{2}) Latency, and (\rom{3}).  Staleness. While CRDTs
try to optimize for latency and staleness, they require a significant
amount of development and verification effort to be expended to ensure
convergence~\cite{kleppmann2017}. In contrast, \quark lets developers
derive convergent-by-construction RDTs from ordinary data data types
that are optimized for latency, but incur a staleness overhead that
delays the time to convergence.
